Title Goes Here

### 2020-04-10 analysis/paper/paper.Rmd
create 3d pca
tried plotly package
what to write

- What are our research questions?

- What is our archaeological background?

- What is our conceptual framework?

- What are our methods?

- What are our results?

- What is the answer to our questions??

@@

@@

>>>>

1. What are our research questions?
- Understanding technological transition during the Late Pleistocene in Korea base on Cultural transmission. 
- What are the social contexts led to the technological transition

2. What is our archaeological background?
-SP was started in Korean Peninsula -> spread to Japan
-Why SP important, Why we care
-Current explantion of SP dispersal: migration vs in situ, in between
 The evidence of the two main models: proportion of SP in assemblage, Y chromosome (two groups of people), composition of assemblage (blade to flake ratio)
 -> no explanation about any mechanism/process/driver for the new technology
 -> lack of details of the lithic technology, human behavior

3. What is our conceptual framework?
- Evolutionary concepts that links stone artifacts to human behavior
- Cultural Evolution : the idea that cultural change constitutes an evolutionary process
- Same general concept of biological evolution: genetic change in population that is inherited over generations. Apply this concept to artifacts or cultural phenomena
- In this paper, I will use Cultural Transmission 
- Cultural Transmssion: way that people learn and pass on information
- Transmission biases
- similar study (CT in stone artifacts) : Mesoudi&O'Brien, Bettinger&Eerkens, Garvey

4. What are our methods?
- Framework: Bettinger&Eerkens
- Two models: Guided variation and Model-based bias =>  Multi-point Origin (in situ) and Single-point Origin (migration model)
- Test by Correlation and Variation (Pearson's r, PCA, Coefficient of Variation) 
- data collection, sites, artifact #
- attributes: landmarks (tps software)

5. What are our results? 
- overall change through time, grouping (chronology, typology)
- correlation
- PCA
- CV

6. What is the answer to our questions??
- low correlation + high variation? High correlation + low variation?
- How was SP spread in Korea, did they have single orgin or multiple origins?

7. What is limitation of this study?
- Maybe applying CV for SP is not ideal. It can be depending on individual knapper's skill and raw materials
- CV: lithis vs pottery

@@

PCA for 3 factors
```{r}
library(plot3D)

scatter3D(pca_col$PC1, pca_col$PC2, pca_col$PC3, bty = "g", pch = 18, col = gg.col(100), 
          main = "PCA", xlab = "PC1",
          ylab = "PC2", zlab = "PC3")

library(plotly)
plot_ly(pca_col, x = ~PC1, y = ~PC2, z = ~PC3, colors = "Dark2", size = I(120)) %>%  add_markers(color = ~phase)

@@

PCA with corrplot package to understand the individual variables
```{r}
library(factoextra) #http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/

fviz_pca_var(PCA, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"))
